### 2024-08-02 getting-started.qmd
20240802 - move player IDs earlier so they can be used for later code
20240802 - actual stats

The `nfl_actualStats_offense_career`, `nfl_actualStats_defense_career`, and `nfl_actualStats_kicking_career` objects are in `player` form.
That is, each row should be uniquely identified by the combination of `player_id`.
Let's rearrange the data accordingly:

@@

nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
  dplyr::bind_rows()
nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
  dplyr::bind_rows()
nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
  dplyr::bind_rows()
```

@@

player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
  dplyr::left_join(
    years_of_experience_offense[,c("player_id","season","years_of_experience")],
    by = c("player_id","season")
  )

# Arrange data
player_stats_weekly_offense <- player_stats_weekly_offense %>% 
  arrange(player_display_name, season, week)

player_stats_weekly_defense <- player_stats_weekly_defense %>% 
  arrange(player_display_name, season, week)

player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
  arrange(player_display_name, season, week)
```

@@

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  dplyr::left_join(
    years_of_experience_offense[,c("player_id","season","years_of_experience")],
    by = c("player_id","season")
  )

# Arrange data
player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  relocate(season, .after = player_id) %>% 
  arrange(player_display_name, season)

player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
  relocate(season, .after = player_id) %>% 
  arrange(player_display_name, season)

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  relocate(season, .after = player_id) %>% 
  arrange(player_display_name, season)
```

>>>>

The `nfl_actualStats_career` objects are in `player` form.
That is, each row should be uniquely identified by the combination of `player_id`.
Let's rearrange the data accordingly:

@@

nfl_actualStats_offense_seasonal_raw <- nfl_actualStats_offense_seasonalList %>% 
  dplyr::bind_rows()
nfl_actualStats_defense_seasonal_raw <- nfl_actualStats_defense_seasonalList %>% 
  dplyr::bind_rows()
nfl_actualStats_kicking_seasonal_raw <- nfl_actualStats_kicking_seasonalList %>% 
  dplyr::bind_rows()
```

save(
  nfl_actualStats_offense_seasonal_raw, nfl_actualStats_defense_seasonal_raw, nfl_actualStats_kicking_seasonal_raw,
  file = "./data/nfl_actualStats_seasonal_raw.RData"
)
```

load("./data/nfl_actualStats_seasonal_raw.RData")
```

The `nfl_actualStats_seasonal` objects are in `player`-`season` form.
That is, each row should be uniquely identified by the combination of `player_id` and `season`.
Let's rearrange the data accordingly:

```{r}
nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)
```

```{r}
nfl_actualStats_offense_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_defense_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_kicking_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```

```{r}
# Sum statistics across a player's years with multiple teams
nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
  group_by(player_id, season) %>% 
  summarise(
    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )

nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
  group_by(player_id, season) %>% 
  summarise(
    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )

# Re-calculate percentage stats
nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
  mutate(
    fg_pct = fg_made / fg_att,
    pat_pct = pat_made / pat_att
  )

# Merge data back with player info
nfl_actualStats_defense_seasonal_raw_playerInfo <- nfl_actualStats_defense_seasonal_raw %>% 
  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
  unique() %>% 
  group_by(player_id) %>% 
  slice_head(n = 1) %>% 
  ungroup()

nfl_actualStats_kicking_seasonal_raw_playerInfo <- nfl_actualStats_kicking_seasonal_raw %>% 
  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
  unique() %>% 
  group_by(player_id) %>% 
  slice_head(n = 1) %>% 
  ungroup()

nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw_playerInfo %>% 
  right_join(
    nfl_actualStats_defense_seasonal,
    by = c("player_id","season")
  )

nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw_playerInfo %>% 
  right_join(
    nfl_actualStats_kicking_seasonal,
    by = c("player_id","season")
  )

# Rearrange data
nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)
```

```{r}
nfl_actualStats_offense_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_defense_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_kicking_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```

@@

player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
  dplyr::left_join(
    years_of_experience_offense[,c("player_id","season","years_of_experience")],
    by = c("player_id","season")
  )
```

The `player_stats_weekly` objects are in `player`-`season`-`week` form.
That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
Let's rearrange the data accordingly:

```{r}
player_stats_weekly_offense <- player_stats_weekly_offense %>% 
  arrange(player_display_name, player_id, season, week)

player_stats_weekly_defense <- player_stats_weekly_defense %>% 
  arrange(player_display_name, player_id, season, week)

player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
  arrange(player_display_name, player_id, season, week)
```

```{r}
player_stats_weekly_offense %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()

player_stats_weekly_defense %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head() #todo: fix duplicate instances

player_stats_weekly_kicking %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head() #todo: fix duplicate instances
```

@@

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  dplyr::left_join(
    years_of_experience_offense[,c("player_id","season","years_of_experience")],
    by = c("player_id","season")
  )
```

The `player_stats_seasonal` objects are in `player`-`season` form.
That is, each row should be uniquely identified by the combination of `player_id` and `season`.
Let's rearrange the data accordingly:

```{r}
player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)
```

Let's check for duplicate `player`-`season` instances:

```{r}
player_stats_seasonal_offense %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

player_stats_seasonal_defense %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

player_stats_seasonal_kicking %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```
